Interval Probability for Fuzzy Quantum Theories

### 2016-08-05 prob.tex
prob.tex Fix small things...

As motivated above, an event $E_{1}$ may have an interval of probability
$[l_{1},r_{1}]$. Assume that another disjoint event $E_{2}$ has
interval probability $[l_{2},r_{2}]$, what is the interval probability
of the event $E_{1}\cup E_{2}$? The answer is somewhat subtle: although
it is possible to use the sum of the intervals $[l_{1}+l_{2},r_{1}+r_{2}]$
as the combined probability, one can do find a much tighter interval
if information \emph{against} the event (i.e., information about the
complement event) is also taken into consideration. Formally, for
a general event $E$ with probability $[l,r]$, the evidence that
contradicts $E$ is an evidence supporting the complement of $E$.
The complement of $E$ must therefore have probability $\left[1-r,1-l\right]$
which we abbreviate $1\tensor*[_{\mathscr{I}}]{-}{}\left[l,r\right]$,
where the preposing $\mathscr{I}$ specifies we subtracts intervals.
Given a collection of intervals $\mathscr{I}$, an $\mathscr{I}$\textendash interval-valued
probability measure is a function $\bar{\mu}:\events\rightarrow\mathscr{I}$
such that~\cite{JamisonLodwick2004}:\footnote{The right-end of the interval-valued probability measure is a special
case of (sub-additive) capacity~\cite{Choquet1954,Graf1980,Goodman2013}.
The left-end of the interval-valued probability measure is a special
case of game~\cite{Shapley1965}.} 
\begin{itemize}
\item $\bar{\mu}(\emptyset)=[0,0]$, 
\item $\bar{\mu}(\Omega)=[1,1]$, 
\item for any event $E$, $\bar{\mu}\left(\Omega\backslash
E\right)=1\tensor*[_{\mathscr{I}}]{-}{}\bar{\mu}\left(E\right)$, and 
\item for a collection $E_{i}$ of pairwise disjoint events, we have 
\begin{equation}
\bar{\mu}\left(\bigcup_{i}E_{i}\right)\subseteq\tensor*[_{\mathscr{I}}]{\sum}{_{i}}\bar{\mu}\left(E_{i}\right)\textrm{ ,}\label{eq:subsetInterval-valuedProbabilityMeasure}
\end{equation}
where $\tensor*[_{\mathscr{I}}]{\sum}{_{i}}[l_{i},r_{i}]=\left[\tensor*[_{\mathbb{R}}]{\sum}{_{i}}l_{i},\tensor*[_{\mathbb{R}}]{\sum}{_{i}}r_{i}\right]$.
We may drop the preposing $\mathscr{I}$ when summands are clearly
intervals. 
\end{itemize}

@@

We will explain why equation (\ref{eq:subsetInterval-valuedProbabilityMeasure})
is expressed using $\subseteq$ by a small example.

@@

\begin{definition}~Given an interval-valued probability measure~$\bar{\mu}:\events\rightarrow\mathscr{I}$:
\begin{itemize}
\item A \emph{core}\footnote{The core is also called a structure~\cite{Weichselberger2000}.}
of $\bar{\mu}$ is the set~\cite{GilboaSchmeidler1994,Marinacci1999,Teran2014}
\[
\mathrm{core}\left(\bar{\mu}\right)=\left\{ \pmeas:\events\rightarrow[0,1]\textrm{ probability measure}\middle|\pmeas\left(E\right)\in\bar{\mu}\left(E\right)\forall E\in\events\right\} \textrm{ .}
\]
\item $\bar{\mu}$ is called \emph{convex} or \emph{2-monotone} if $\bar{\mu}\left(E_{1}\cup E_{2}\right)\tensor*[_{\mathscr{I}}]{+}{}\bar{\mu}\left(E_{1}\cap E_{2}\right)\subseteq\bar{\mu}\left(E_{1}\right)\tensor*[_{\mathscr{I}}]{+}{}\bar{\mu}\left(E_{2}\right)$
for all $E_{1},E_{2}\in\events$~\cite{Shapley1965,GilboaSchmeidler1994,Marinacci1999,Teran2014}.\footnote{\yutsung{If we want to use convex, we need to explain the physical
meaning of convex interval-valued probability measures...? If there
is a good explanation, we may only consider convex interval-valued
probability measures, and replace the definition in the beginning...?}}
\end{itemize}
\qed\end{definition}

@@

\begin{definition}~Given a quantum interval-valued probability measure~$\bar{\mu}:\events\rightarrow\mathscr{I}$:
\begin{itemize}
\item A \emph{core} of $\bar{\mu}$ is the set 
\[
\mathrm{core}\left(\bar{\mu}\right)=\left\{ \mu:\events\rightarrow[0,1]\textrm{ quantum probability measure}\middle|\mu\left(P\right)\in\bar{\mu}\left(P\right)\forall P\in\events\right\} \textrm{ .}
\]
\item $\bar{\mu}$ is called convex if 
\begin{equation}
\bar{\mu}\left(P_{1}\tensor*[_{\mathscr{O}}]{\vee}{}P_{2}\right)\tensor*[_{\mathscr{I}}]{+}{}\bar{\mu}\left(P_{1}P_{2}\right)\subseteq\bar{\mu}\left(P_{1}\right)\tensor*[_{\mathscr{I}}]{+}{}\bar{\mu}\left(P_{2}\right)\label{eq:QuantumInterval-valuedProbability-Convex}
\end{equation}
 for all commuting $P_{1},P_{2}\in\events$. 
\end{itemize}
\qed\end{definition}

@@

Next, we need to verify that $\bar{\mu}$ is convex. First, for orthogonal
projections $P_{1}$ and $P_{2}$, equation (\ref{eq:QuantumInterval-valuedProbability-Inclusion})
implies equation (\ref{eq:QuantumInterval-valuedProbability-Convex}).
Second, when $P_{1}=P_{2}+P_{3}$, equation (\ref{eq:QuantumInterval-valuedProbability-Convex})
holds trivially. Hence, we only need to verify the case which $P_{1}=\proj{\psi_{1}}\tensor*[_{\mathscr{O}}]{+}{}\proj{\psi_{3}}$
and $P_{2}=\proj{\psi_{2}}\tensor*[_{\mathscr{O}}]{+}{}\proj{\psi_{3}}$,
where $\left\{ \ket{\psi_{1}},\ket{\psi_{2}},\ket{\psi_{3}}\right\} $
is an orthonormal basis. Because we have 
\begin{eqnarray*}
\bar{\mu}\left(P_{1}\tensor*[_{\mathscr{O}}]{\vee}{}P_{2}\right) & = & \bar{\mu}\left(\mathbb{1}\right)=\necess\\
\bar{\mu}\left(P_{1}P_{2}\right) & = & \bar{\mu}\left(\proj{\psi_{3}}\right)\subseteq\unknown
\end{eqnarray*}
and $\left\{ \bar{\mu}\left(P_{1}\right),\bar{\mu}\left(P_{2}\right)\right\} $
is either $\left\{ \bar{\mu}\left(\necess\right),\bar{\mu}\left(\unknown\right)\right\} $
or $\left\{ \bar{\mu}\left(\unknown\right),\bar{\mu}\left(\unknown\right)\right\} $.
Therefore, we have
\[
\bar{\mu}\left(P_{1}\tensor*[_{\mathscr{O}}]{\vee}{}P_{2}\right)\tensor*[_{\mathscr{I}}]{+}{}\bar{\mu}\left(P_{1}P_{2}\right)\subseteq\bar{\mu}\left(P_{1}\right)\tensor*[_{\mathscr{I}}]{+}{}\bar{\mu}\left(P_{2}\right)\textrm{ .}
\]

$\bar{\mu}$ cannot correspond to any quantum probability measure.
Suppose $\bar{\mu}(P)=\iota\left(\mu(P)\right)$ for some $\mu:\events\rightarrow\left[0,1\right]$.
We must have 
\begin{equation}
\mu(\proj{0})=\mu(\proj{\ps})=0\textrm{ .}\label{eq:probability-zero-on-states}
\end{equation}
By Gleason's theorem, there is a mixed state~$\rho=\sum_{j}q_{j}\proj{\phi_{j}}$
such that $\mu\left(P\right)=\sum_{j=1}^{N}q_{j}\ip{\phi_{j}}{P\phi_{j}}$,
where $\sum_{j}q_{j}=1$ and $q_{j}>0$. The only pure state~$\ket{\phi}$
satisfies
\[
\ip{\phi}{0}=\ip{\phi}{\ps}=0
\]
is $\ket{2}$, but clearly $\bar{\mu}(P)\ne\iota\left(\mu_{\ket{2}}^{B}(P)\right)$.\qed\end{example}

>>>>

As motivated above, an event $E_{1}$ may have an interval of probability
$[l_{1},r_{1}]$. Assume that another disjoint event $E_{2}$ has
interval probability $[l_{2},r_{2}]$, what is the interval probability
of the event $E_{1}\cup E_{2}$? The answer is somewhat subtle: although
it is possible to use the sum of the intervals $[l_{1}+l_{2},r_{1}+r_{2}]$
as the combined probability, one can do find a much tighter interval
if information \emph{against} the event (i.e., information about the
complement event) is also taken into consideration. Formally, for
a general event $E$ with probability $[l,r]$, the evidence that
contradicts $E$ is an evidence supporting the complement of $E$.
The complement of $E$ must therefore have probability $\left[1-r,1-l\right]$
which we abbreviate $1\tensor*[_{\mathscr{I}}]{-}{}\left[l,r\right]$,
where the preposing $\mathscr{I}$ specifies we subtracts intervals.
Given a collection of intervals $\mathscr{I}$, an $\mathscr{I}$\textendash interval-valued
probability measure is a function $\bar{\mu}:\events\rightarrow\mathscr{I}$
such that~\cite{JamisonLodwick2004}:\footnote{The right-end of the interval-valued probability measure is a special
case of (sub-additive) capacity~\cite{Choquet1954,Graf1980,Goodman2013}.
The left-end of the interval-valued probability measure is a special
case of game~\cite{Shapley1965}.} 
\begin{itemize}
\item $\bar{\mu}(\emptyset)=[0,0]$, 
\item $\bar{\mu}(\Omega)=[1,1]$, 
\item for any event $E$, $\bar{\mu}\left(\Omega\backslash
E\right)=1\tensor*[_{\mathscr{I}}]{-}{}\bar{\mu}\left(E\right)$, and 
\item for a collection $E_{i}$ of pairwise disjoint events, we have $\bar{\mu}\left(\bigcup_{i}E_{i}\right)\subseteq\tensor*[_{\mathscr{I}}]{\sum}{_{i}}\bar{\mu}\left(E_{i}\right)$,
where $\tensor*[_{\mathscr{I}}]{\sum}{_{i}}[l_{i},r_{i}]=\left[\tensor*[_{\mathbb{R}}]{\sum}{_{i}}l_{i},\tensor*[_{\mathbb{R}}]{\sum}{_{i}}r_{i}\right]$.
We may drop the preposing $\mathscr{I}$ when summands are clearly
intervals. 
\end{itemize}

@@

\noindent We will explain why the last condition is expressed using
$\subseteq$ by a small example.

@@

\begin{definition}~Given an interval-valued probability measure~$\bar{\mu}:\events\rightarrow\mathscr{I}$:
\begin{itemize}
\item A \emph{core}\footnote{The core is also called a structure~\cite{Weichselberger2000}.}
of $\bar{\mu}$ is the set $\mathrm{core}\left(\bar{\mu}\right)=\left\{ \pmeas:\events\rightarrow[0,1]\textrm{ probability measure}\middle|\pmeas\left(E\right)\in\bar{\mu}\left(E\right)\forall E\in\events\right\} $
\cite{GilboaSchmeidler1994,Marinacci1999,Teran2014}.
\item $\bar{\mu}$ is called \emph{convex} or \emph{2-monotone} if $\bar{\mu}\left(E_{1}\cup E_{2}\right)\tensor*[_{\mathscr{I}}]{+}{}\bar{\mu}\left(E_{1}\cap E_{2}\right)\subseteq\bar{\mu}\left(E_{1}\right)\tensor*[_{\mathscr{I}}]{+}{}\bar{\mu}\left(E_{2}\right)$
for all $E_{1},E_{2}\in\events$~\cite{Shapley1965,GilboaSchmeidler1994,Marinacci1999,Teran2014}.\footnote{\yutsung{If we want to use convex, we need to explain the physical
meaning of convex interval-valued probability measures...? If there
is a good explanation, we may only consider convex interval-valued
probability measures, and replace the definition in the beginning...?}}
\end{itemize}
\qed\end{definition}

@@

\begin{definition}~Given a quantum interval-valued probability measure~$\bar{\mu}:\events\rightarrow\mathscr{I}$:
\begin{itemize}
\item A \emph{core} of $\bar{\mu}$ is the set $\mathrm{core}\left(\bar{\mu}\right)=\left\{ \mu:\events\rightarrow[0,1]\textrm{ quantum probability measure}\middle|\mu\left(P\right)\in\bar{\mu}\left(P\right)\forall P\in\events\right\} $.
\item $\bar{\mu}$ is called convex if 
\begin{equation}
\bar{\mu}\left(P_{1}\tensor*[_{\mathscr{O}}]{\vee}{}P_{2}\right)\tensor*[_{\mathscr{I}}]{+}{}\bar{\mu}\left(P_{1}P_{2}\right)\subseteq\bar{\mu}\left(P_{1}\right)\tensor*[_{\mathscr{I}}]{+}{}\bar{\mu}\left(P_{2}\right)\label{eq:QuantumInterval-valuedProbability-Convex}
\end{equation}
 for all commuting $P_{1},P_{2}\in\events$. 
\end{itemize}
\qed\end{definition}

@@

Next, we need to verify that $\bar{\mu}$ is convex. First, for orthogonal
projections $P_{1}$ and $P_{2}$, equation (\ref{eq:QuantumInterval-valuedProbability-Inclusion})
implies equation (\ref{eq:QuantumInterval-valuedProbability-Convex}).
Second, when $P_{1}=P_{2}+P_{3}$, equation (\ref{eq:QuantumInterval-valuedProbability-Convex})
holds trivially. Hence, we only need to verify the case which $P_{1}=\proj{\psi_{1}}\tensor*[_{\mathscr{O}}]{+}{}\proj{\psi_{3}}$
and $P_{2}=\proj{\psi_{2}}\tensor*[_{\mathscr{O}}]{+}{}\proj{\psi_{3}}$,
where $\left\{ \ket{\psi_{1}},\ket{\psi_{2}},\ket{\psi_{3}}\right\} $
is an orthonormal basis. Because we have 
\begin{eqnarray*}
\bar{\mu}\left(P_{1}\tensor*[_{\mathscr{O}}]{\vee}{}P_{2}\right) & = & \bar{\mu}\left(\mathbb{1}\right)=\necess\\
\bar{\mu}\left(P_{1}P_{2}\right) & = & \bar{\mu}\left(\proj{\psi_{3}}\right)\subseteq\unknown
\end{eqnarray*}
and $\bar{\mu}\left(P_{1}\right)\tensor*[_{\mathscr{I}}]{+}{}\bar{\mu}\left(P_{2}\right)$
is either $\necess\tensor*[_{\mathscr{I}}]{+}{}\unknown$ or $\unknown\tensor*[_{\mathscr{I}}]{+}{}\unknown$.
Therefore, we have$\bar{\mu}\left(P_{1}\tensor*[_{\mathscr{O}}]{\vee}{}P_{2}\right)\tensor*[_{\mathscr{I}}]{+}{}\bar{\mu}\left(P_{1}P_{2}\right)\subseteq\bar{\mu}\left(P_{1}\right)\tensor*[_{\mathscr{I}}]{+}{}\bar{\mu}\left(P_{2}\right)$.

$\bar{\mu}$ cannot correspond to any quantum probability measure.
Suppose $\mu(P)\in\bar{\mu}(P)$ for some $\mu:\events\rightarrow\left[0,1\right]$.
We must have 
\begin{equation}
\mu(\proj{0})=\mu(\proj{\ps})=0\textrm{ .}\label{eq:probability-zero-on-states}
\end{equation}
By Gleason's theorem, there is a mixed state~$\rho=\sum_{j}q_{j}\proj{\phi_{j}}$
such that $\mu\left(P\right)=\sum_{j=1}^{N}q_{j}\ip{\phi_{j}}{P\phi_{j}}$,
where $\sum_{j}q_{j}=1$ and $q_{j}>0$. The only pure state~$\ket{\phi}$
satisfies 
\[
\ip{\phi}{0}=\ip{\phi}{\ps}=0
\]
is $\ket{2}$, but clearly $\bar{\mu}(P)\ne\iota\left(\mu_{\ket{2}}^{B}(P)\right)$.\qed\end{example}
